---
title: "dplyr"
output: html_notebook
---
dplyr quick notes
Main player of tidying data for the tidyverse. 

6 main functions
filter()
arrange()
select()
mutate()
group_by()
summarize()

simple use 
pull()
n()/count()
glimpse()

First, get your data set as a tibble
```{r}
mtcars <- as_tibble(mtcars)

mtcars
```
Just like in python, you can use head and tail functions to limit down the dataset. 

Structure is largely the same. 

```{r}
head(mtcars, 5)
#tail(mtcars, 5)
```

Can also use the slice function (similar to .loc in pandas). Starts with 1 instead of 0 for python.

```{r}
slice(mtcars, 1,3,5)
```
Glimpse can be used to pull a quick visualization of the dataset. 

```{r}
glimpse(mtcars)
```
Filter is a crucial dplyr function.  It has to have == to work, single = is read as a variable assignment which is not what you want. 

simply use a , followed by a column name to filter your dataset down. 

To view all 8 cylinder vehicles from mtcars...

```{r}
#filter(mtcars, cyl == 8)



#Can also use more complex features. 
mtcars %>%
  filter(cyl %in% c(4,6))
#Excludes everything that does not have 4 or 6 cylinders. 



#NOT EQUAL TO
mtcars %>%
  filter(cyl !=4)
#Shows everything that is not 4 cylinder. 
```
Arrange - Sorts the dataset top to bottom. 

arrange will default as a low to high, use desc to sort high to low

```{r}
#Ascending
mtcars %>%
  arrange(mpg)

#Descending
mtcars %>%
  arrange(desc(mpg))

#Can stack multiple variables....
mtcars %>%
  arrange(desc(cyl), desc(hp))
```

Combining layers together.......

```{r}
mtcars %>%
  select(cyl,mpg) %>%
  glimpse()

#Selecting multiple columns
mtcars %>%
  select(cyl,mpg, hp, disp) %>%
  glimpse()

#Dropping columns from the select query. 
mtcars %>%
  select(disp, cyl, hp, mpg) %>%
  select(-mpg) %>%       #This drops mpg
  glimpse()

#Selecting columns that begin with something
mtcars %>%
  select(starts_with('c'), starts_with('h')) %>%
  glimpse()

```
Mutate to transform the data table for manipulation

mutates the cyl column to move from a integer to a factor (which is a string)
```{r}
mtcars %>%
  mutate(cyl = factor(cyl,
                      levels = c(4,6,8),
                      labels = c('4 Cyl','6 Cyl','8 Cyl'))) %>%
  select (cyl) %>%
  glimpse()

```
Can also use mutations to change and modify datasets for different units

```{r}
mtcars %>%
  mutate(miles_per_liter = mpg * 3.78541,
         miles_per_gallon = miles_per_liter / 3.78541) %>%
  select(miles_per_liter, mpg, miles_per_gallon) %>%
  head(10)

mtcars %>%
  group_by(cyl) %>%
  #slice(1) #Grabs the first observation for each of the cyl variables
  #filter(hp == max(hp))  #max hp observed
  #top_n(1, hp)  #top observation by column
  #arrange(desc(hp))
  #count()

```

aggregation functions
```{r}
mtcars %>%
  summarize(mean=mean(mpg))

#why not...
# mtcars %>% mean(mpg)  This looks for a vector. 

mtcars %>%
  pull(mpg) %>%
  mean()  #This will run the mean on the pull, see below for result.


mtcars %>%
  summarize(median(mpg))

mtcars %>%
  summarize(median_mpg = mean(mpg))

```
Summarize gets really powerful when combined with GROUP_BY

```{r}
mtcars %>%
  group_by(cyl) %>%
  summarize(mean_mpg = mean(mpg),
            sd_mpg =sd(mpg),
            n = n())

```
Mutate_at()
```{r}
mtcars %>%
  mutate_at(.vars = vars(cyl, am), factor) %>%
  select(cyl, am) %>%
  glimpse()

mtcars %>%
  group_by(cyl) %>%
  summarize_at (.vars = vars(mpg, disp, qsec),.funs=mean)

```










