---
title: "ggplot2"
output: html_notebook
---
Could discuss this for hours, there's a billion things you can do with these. 
Great for exploratory data analysis. 
Final step after data is tidyied and dplyr'd. 

########################################################

3 core parts

ggplot - builds the base layer

geom_ is the shape
- geom_point()
- geom_line()
- geom_bar()
- geom_boxplot()
- geom_?()

aes is the mapping or relationship of the data
- Horizontal dimensions (x)
- vertical dimensions (y)
- color
- Shape
- Size
- Transparency
- Relationship


Basic ggplot call
```{r}
#ggplot(data = <DATA>, mapping = aes(<MAPPINGS>)) +
 # <GEOM_FUNCTION>()

mtcars <- as_tibble(mtcars)

ggplot(mtcars, aes(x = cyl, y = mpg)) +
  geom_point()

#ggplot generates below in the code. 
```
So many types of graphs....
- geom_point() - Scatter plot
- geom_dotplot() - Dot Plot
- geom_hline() - Horizontal Ref Line
- geom_vline() - Vertical refernce line
- geom_boxplot() - box error plot
- geom_density() - smoothed density estimates
- geom_errorbarh() - horizontal error bars
- geom_hex() - Hexagonal heatmap of 2d bin counts
- geom_jitter() - Jittered points
- geom_linearange() - Vertical interval line
- geom_pointrange() - Vertical point line
- geom_line() - connect observations line
- geom_step() - connect observations via step lines
- geom_polygon() - Polygons
- geom_segment() - line segment
- geom_ribbon() - Ribbon plot
- geom_area() - Area plot
- geom_rug() - rug plots in the margins
- geom_smooth() - smoothed conditional means
- geom_label() - label points with text
- geom_text() - add text
- geom_violin() - Violin plot
- geom_sf() - Visual sf plot
- geom_map() - Plot map
- geom_qq_line() - a Quantile-quantile plot
- geom_histogram() - Histogram plot

```{r}
install.packages('gapminder')

library('gapminder')
gapminder_df <- as_tibble(gapminder)

gapminder_df %>%
  filter(country == 'Turkey') %>%
  #filter(lifeExp < 70) %>%
  ggplot(aes(x = year, y = lifeExp, colour = country)) +
  geom_line(size = 1)

```

```{r}
bar_df <-gapminder_df %>%
  filter(year == 2007 & continent == 'Americas') %>%
  arrange(desc(lifeExp)) %>%
  head(8)

#A different way, assigning a plot to a variable for future calls, can reduce code. 
(bars <-ggplot(bar_df, aes(x = country, y = lifeExp)) +
    geom_bar(stat = "identity"))  #can also use, geom_col()) instead of geom_bar

#Histograms
hist_plot <-gapminder_df %>%
  filter(year == 2007) %>%
  ggplot(aes(x = lifeExp)) +
  geom_histogram(binwidth = 5, color = "White")

prettify<-hist_plot +
            labs(x = 'Life Expectancy (Years)',
                 y = 'Count',
                 title = "Life Expectancy",
                 subtitle = 'Year = 2007',
                 caption = "Source: Gapminder")
prettify

#To save plots as .png
ggsave('life_exp_2007.png',prettify,
       height = 4, width = 6, units = 'in', dpi = 450)

```
Using themes to save code in GGPLOT2

```{r}
install.packages("ggthemes")

gapminder_df %>%
  filter(country == 'China' | country == 'United States') %>%
  ggplot(aes(x = year, y = lifeExp, color = country)) +
  geom_line(size = 1) +
  ggthemes::theme_few()

#theme_minimal()
#ggthemes::theme_economist_white()
#theme_bw()
#theme_classic()
#ggthemes::theme_few()
```











